### 2019-05-03 initial/notes.tex
maybe?

In the presence of tides, and further assuming $s \ll l_1$, we may write (Lai
2012, Equations 43--44, also Ward 1975 Equation 9 \& 13)
\begin{align}
    \frac{1}{s}\rd{s}{t}
        = \frac{1}{s}\rd{s}{t}
        &= \frac{1}{t_s}\frac{L}{2S}\s*{\cos\theta
            - \frac{s}{2\Omega_1}\p*{1 + \cos^2\theta}},\\
    \rd{\theta}{t} &= -\frac{1}{t_s}\frac{L}{2S}
        \sin\theta\p*{1 - \frac{s}{2\Omega_1}\cos\theta}.
\end{align}

It is perhaps easiest to define $\frac{s}{s_c} =
\frac{\omega_{s1}}{\omega_{1p}\cos I}$ and $\epsilon = \frac{L}{2St_s
\omega_{1p}\cos I}$ while rescaling time $\tau = \omega_{1p}\cos(I) t$, so that
we obtain equations of motion
\begin{align}
    \rd{\hat{s}}{\tau}
        &= \frac{s}{s_c}\p*{\hat{s} \cdot \hat{l}_1}\p*{\hat{s} \times \hat{l}_1}
            - \hat{s} \times \hat{l}_p
            + \epsilon \p*{1 - \frac{s}{2\Omega_1}\p*{\hat{l}_1 \cdot \hat{s}}}
                \hat{s} \times \p*{\hat{l}_1 \times \hat{s}},\\
    \rd{s}{\tau}
        &= \epsilon s \p*{\hat{s} \cdot \hat{l}_1 - \frac{s}{2\Omega_1}\p*{1
            + \p*{\hat{s} \cdot \hat{l}_1}^2}}.
\end{align}
$s_c$ has the interpretation of being the critical spin such that the $s1$
coupling is roughly equal strength to the $1p$ coupling. There then seem to be a
few outcomes that we might expect:
\begin{itemize}
    \item Fast evolution towards CS1, then tides will slowly change $s$ without
        changing $\hat{s}$.

@@

To make the equations more anemable to analytic analysis (not simulation), let's
write down the EOM in $(\mu, \phi)$ coordinates again. The $\phi$ EOM does not
change from the tide-free case, so we can reuse earlier equation:
\begin{align}
    \pd{\phi}{\tau} &= \frac{s}{s_c}\mu
        - \p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi} ,\\
    \pd{\mu}{\tau} &= -\sin I \sin \phi +
        \epsilon \p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\\
    \rd{s}{\tau}
        &= \epsilon s \p*{\mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2}}.
\end{align}
Assuming $s \gg s_c$ the strong spin-orbit coupling regime, let's first try
assuming $\mu$ is roughly constant over the course of a precession period, then
we can average out the $\phi$ dependencies. Then $\phi$ drops out of the EOM,
and we have approximate averaged equations
\begin{align}
    \pd{\mu}{(\epsilon\tau)} &\approx \p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\nonumber\\
        &\approx \p*{1 - \mu^2}
            \p*{- \frac{s}{2\Omega_1}\mu},\\
    \frac{1}{s}\rd{s}{(\epsilon\tau)}
        &\approx \mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2},\nonumber\\
        &\approx - \frac{s}{2\Omega_1}.
\end{align}
In the last term, we note $s \gtrsim 2\Omega_1$ initially, while $\mu \leq 1$,
so we drop both the linear contribution from $\mu$ and approximate $\p*{1 +
\mu^2} \approx 1$.

Making all these approximations, we identify approximate analytic solution
$s(\epsilon \tau) \approx \frac{2\Omega_1}{\epsilon \tau + C}$ with constant of
integration $C$ set by initial conditions, or more precisely
\begin{equation}
    s(\epsilon \tau) \approx \frac{2\Omega_1}{\epsilon \tau +
        \frac{2\Omega_1}{s(0)}}.
\end{equation}
The critical timescale is clearly $\epsilon \tau \sim 1$ (since $2\Omega_1/s(0)
\ll 1$) at which $\frac{s}{2\Omega_1} \sim 1$ and synchronization sets in.

The other timescale of interest is how long it takes $\hat{s}$ to reach a
Cassini State. We may present crude estimates in the $s \gg s_c$ strong
spin-orbit coupling limit; consider starting $\mu = -1 + \delta$, then to
leading order
\begin{align}
    \pd{\delta}{(\epsilon\tau)} &\approx 2\delta \frac{s}{2\Omega_1},\nonumber\\
    \delta(\epsilon\tau) &\approx
        \delta_0\exp\s*{\frac{\epsilon \tau}{s/\Omega_1}}.
\end{align}
Setting $\delta \sim 1$ gives $\p*{\epsilon\tau}_{CS} \sim
-\frac{\Omega_1}{s}\ln \delta_0$ the timescale it takes to reach $\mu = 0$ which
is of order the timescale to reach the CS\@. This tells us that at displacements
of $\delta_0 \gtrsim e^{-s/\Omega_1}$ is where states will be able to reach
Cassini States. So for any respectable $s/\Omega_1$, the CS timescale will be
short compared to the synchronization timescale, so we should expect the
adiabatic evolution near CS1, CS2 as $\frac{s}{s_c}$ changes governs the
long-term observed states. Specifically, states that end up near CS1 should
(slowly!) hit the resonant obliquity excitation that Kassandra studied, while
states near CS2 should stay on CS2 as $\eta$ grows past $\eta_{crit}$, since it
is far from bifurcation.

>>>>

In the presence of tides, and further assuming $s \ll l_1$, we may write (Lai
2012, Equations 43--44, also Ward 1975 Equation 9 \& 13)
\begin{align}
    \frac{1}{s}\rd{s}{t}
        = \frac{1}{s}\rd{s}{t}
        &= \frac{1}{t_s}\frac{L}{2S}\s*{\cos\theta
            - \frac{s}{2\Omega_1}\p*{1 + \cos^2\theta}},\\
    \rd{\theta}{t} &= -\frac{1}{t_s}\frac{L}{2S}
        \sin\theta\p*{1 - \frac{s}{2\Omega_1}\cos\theta}.
\end{align}
Note that $L = \mu a^2\Omega_1, S = Is$ are the orbital and spin angular momenta
espectively.

It is perhaps easiest to define $\frac{s}{s_c} =
\frac{\omega_{s1}}{\omega_{1p}\cos I}$ and $\epsilon\frac{\Omega_1}{s} =
\frac{L}{2St_s \omega_{1p}\cos I}$ while rescaling time $\tau =
\omega_{1p}\cos(I) t$, so that we obtain equations of motion
\begin{align}
    \rd{\hat{s}}{\tau}
        &= \frac{s}{s_c}\p*{\hat{s} \cdot \hat{l}_1}\p*{\hat{s} \times \hat{l}_1}
            - \hat{s} \times \hat{l}_p
            + \frac{\epsilon \Omega_1}{s}
                \p*{1 - \frac{s}{2\Omega_1}\p*{\hat{l}_1 \cdot \hat{s}}}
                    \hat{s} \times \p*{\hat{l}_1 \times \hat{s}},\\
    \rd{s}{\tau}
        &= \epsilon \Omega_1 \p*{\hat{s} \cdot \hat{l}_1 - \frac{s}{2\Omega_1}\p*{1
            + \p*{\hat{s} \cdot \hat{l}_1}^2}}.
\end{align}
$s_c$ has the interpretation of being the critical spin such that the $s1$
coupling is roughly equal strength to the $1p$ coupling. There then seem to be a
few outcomes that we might expect:
\begin{itemize}
    \item Fast evolution towards CS1, then tides will slowly change $s$ without
        changing $\hat{s}$.

@@

To make the equations more anemable to analytic analysis (not simulation), let's
write down the EOM in $(\mu, \phi)$ coordinates again. The $\phi$ EOM does not
change from the tide-free case, so we can reuse earlier equation:
\begin{align}
    \pd{\phi}{\tau} &= \frac{s}{s_c}\mu
        - \p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi} ,\\
    \pd{\mu}{\tau} &= -\sin I \sin \phi +
        \epsilon \frac{\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\\
    \rd{s}{\tau}
        &= \epsilon \Omega_1 \p*{\mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2}}.
\end{align}
Assuming $s \gg s_c$ the strong spin-orbit coupling regime, let's first try
assuming $\mu$ is roughly constant over the course of a precession period, then
we can average out the $\phi$ dependencies. Then $\phi$ drops out of the EOM,
and we have approximate averaged equations
\begin{align}
    \pd{\mu}{(\epsilon\tau)} &\approx \frac{\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\nonumber\\
        &\approx \p*{1 - \mu^2}
            \p*{\frac{\Omega_1}{s} - \frac{1}{2}\mu},\\
    \frac{1}{\Omega_1}\rd{s}{(\epsilon\tau)}
        &\approx \mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2},\nonumber\\
        &\approx - \frac{s}{2\Omega_1}.
\end{align}
In the last term, we note $s \gtrsim 2\Omega_1$ initially, while $\mu \leq 1$,
so we drop both the linear contribution from $\mu$ and approximate $\p*{1 +
\mu^2} \approx 1$.

With all these approximations, we clearly obtain $s(\tau) \approx
s(0)e^{-\frac{\epsilon}{2}\tau}$, so the critical synchronization
timescale is $\tau_{sync} \sim \frac{2}{\epsilon}$.

The other timescale of interest is how long it takes $\hat{s}$ to reach a
Cassini State. We may present crude estimates in the $s \gg s_c$ strong
spin-orbit coupling limit; consider starting $\mu = -1 + \delta$, then to
leading order
\begin{align}
    \pd{\delta}{\tau} &\approx +\epsilon \delta,\nonumber\\
    \delta(\epsilon\tau) &\approx \delta_0e^{\epsilon \tau}
\end{align}
Setting $\delta \sim 1$ gives $\tau_{CS} = -\frac{1}{\epsilon}\ln \delta_0$ the
timescale it takes to reach $\mu = 0$ which is of order the timescale to reach
the CS\@. This tells us that at displacements of $\delta_0 \gtrsim
e^{-2}$ is where states will be able to reach Cassini States before
synchronization happens. This reinforces our expectation that synchronization is
probably no faster than reaching the Cassini state, but that a significant
fraction of initial conditions below the separatrix should be able to reach a
Cassini State before tides synchronize.
